---
title: "Code & Data Resources"
format:
  html:
    echo: false
    mainfont: 'inter'
    table-of-contents: true
    number-sections: true
    number-depth: 2
    embed-resources: true
---

```{r}
#| label: 'pre-reqs'
#| include: false
library(duckdb)
library(dplyr)
library(reactable)
library(htmltools)
```

## Script Workflow — `code/weight_contam_flow/`

The scripts in this folder implement a geospatial pipeline that estimates contamination
exposure for US census tracts by combining EPA public water system (PWS) service area
boundaries with census population data and contaminant measurements.

### Pipeline Diagram

```{mermaid}
flowchart TD
    A["<b>Census Decennial API</b><br/>(2010)"]
    B["<b>00_convert_to_geoparquet.R</b><br/>Download tract geometries &amp;<br/>population; convert to Parquet"]
    C["tract10_national.parquet"]

    D["<b>Local Source Files</b><br/>Block group SHPs<br/>EPA Service_Areas_V_3_0.gpkg"]

    E["<b>01_ingest_shapefiles_duckdb.R</b><br/>Contiguous US → EPSG:5070"]
    F["<b>01_ingest_AK_HI_duckdb.R</b><br/>Alaska → EPSG:3338<br/>Hawaii → EPSG:6628"]

    G["<b>DuckDB</b><br/>tract, block group &amp;<br/>EPA boundary tables"]

    H["<b>02_intersect_census_pws.R</b><br/>Spatial intersection of census units<br/>&amp; water system boundaries;<br/>calculates overlap areas"]

    I["<b>DuckDB</b><br/>Intersection tables +<br/>weighting schema"]

    J["<b>PWS_JOINED_BG_POP.csv</b><br/>(Contaminant measurements)"]

    K["<b>03_add_population_weights.R</b><br/>Areal interpolation;<br/>tract-level weight normalization"]

    M["<b>contam_tract_level</b><br/>Weighted contaminant values<br/>at census tract level"]

    A --> B --> C
    D --> E
    D --> F
    C --> E
    C --> F
    E --> G
    F --> G
    G --> H --> I
    I --> K
    J --> K
    K --> M

    style G fill:#dbeafe,stroke:#3b82f6
    style I fill:#dbeafe,stroke:#3b82f6
    style M fill:#dcfce7,stroke:#16a34a
    style C fill:#fef9c3,stroke:#ca8a04
```

### Script Descriptions

```{r}
#| label: 'script-table'
scripts <- tibble::tribble(
  ~Script                                                                                                                                                                                                                                                                              , ~Purpose       ,
  ~`Key Inputs`                                                                                                                                                                                                                                                                        , ~`Key Outputs` ,

  "00_convert_to_geoparquet.R"                                                                                                                                                                                                                                                         ,
  "Downloads 2010 US census tract geometries and total population (P001001) via the tidycensus package, then writes the result to Apache Parquet format for efficient downstream use."                                                                                                 ,
  "Census Decennial API (2010)"                                                                                                                                                                                                                                                        ,
  "data/source_data/tracts/tract10_national.parquet"                                                                                                                                                                                                                                   ,

  "01_ingest_shapefiles_duckdb.R"                                                                                                                                                                                                                                                      ,
  "Reads contiguous US census tracts, block groups, and EPA v3 water system boundaries. Fixes invalid geometries, resolves duplicate PWSIDs via st_union(), reprojects all layers to EPSG:5070, and loads them into DuckDB."                                                           ,
  "tract10_national.parquet · Block group SHPs · Service_Areas_V_3_0.gpkg"                                                                                                                                                                                                             ,
  "DuckDB: tract10_5070 · block10_group_5070 · epa_water_v3_5070"                                                                                                                                                                                                                      ,

  "01_ingest_AK_HI_duckdb.R"                                                                                                                                                                                                                                                           ,
  "Same ingestion workflow as the contiguous script, but filters to Alaska (FIPS 02) and Hawaii (FIPS 15) and reprojects each state to its own CRS to preserve spatial accuracy."                                                                                                      ,
  "Same source files filtered to AK / HI"                                                                                                                                                                                                                                              ,
  "DuckDB: ak_*_3338 (Alaska) · hi_*_6628 (Hawaii)"                                                                                                                                                                                                                                    ,

  "02_intersect_census_pws.R"                                                                                                                                                                                                                                                          ,
  "Performs spatial intersections between census units (tracts and block groups) and EPA water system boundaries for all regions. Calculates overlap areas using ST_Area() and ST_Intersection(), then assembles combined US-wide weighting tables in a dedicated 'weighting' schema." ,
  "DuckDB spatial tables from the two ingest scripts"                                                                                                                                                                                                                                  ,
  "DuckDB: tract/block intersection tables · weighting.us_tract10_pws_v3 · weighting.us_bg10_group_pws_v3"                                                                                                                                                                             ,

  "03_add_population_weights.R"                                                                                                                                                                                                                                                        ,
  "(v1) Implements areal interpolation: distributes block group population proportionally by overlap area, normalizes weights at the tract level when a tract is served by multiple systems, and filters to tracts with >50% public water coverage."                                   ,
  "Intersection tables · PWS_JOINED_BG_POP.csv"                                                                                                                                                                                                                                        ,
  "In-memory: contam_tract_level (weighted contaminant values at tract level)"
)

reactable(
  scripts,
  columns = list(
    Script = colDef(
      name = "Script",
      minWidth = 180,
      style = list(fontFamily = "monospace", fontSize = "0.85em")
    ),
    Purpose = colDef(name = "Purpose", minWidth = 280),
    `Key Inputs` = colDef(name = "Key Inputs", minWidth = 200),
    `Key Outputs` = colDef(name = "Key Outputs", minWidth = 200)
  ),
  striped = TRUE,
  highlight = TRUE,
  bordered = TRUE,
  pagination = FALSE,
  wrap = TRUE,
  theme = reactableTheme(
    headerStyle = list(background = "#f1f5f9", fontWeight = "600")
  )
)
```

---

## Database Schema — `data/tract_water_system_db.duckdb`

The DuckDB database stores all intermediate geospatial and intersection tables produced
by the pipeline. It contains two schemas:

- **`main`** — Base geometry tables and per-region intersection outputs
- **`weighting`** — US-wide combined intersection tables (no geometry) used for population weighting

```{r}
#| label: 'db-connect'
#| include: false
con <- dbConnect(
  duckdb(),
  dbdir = 'C:/git/epa_water_stuff/data/tract_water_system_db.duckdb',
  read_only = TRUE
)
```

### Table Overview

```{r}
#| label: 'table-overview'
table_descriptions <- c(
  "main.tract10_5070" = "Contiguous US census tracts (2010), EPSG:5070",
  "main.block10_group_5070" = "Contiguous US census block groups (2010), EPSG:5070",
  "main.epa_water_v3_5070" = "EPA v3 public water system service area boundaries, EPSG:5070",
  "main.tract10_pws_v3_5070" = "Contiguous US tract × water system spatial intersections with overlap areas, EPSG:5070",
  "main.block10_group_pws_v3_5070" = "Contiguous US block group × water system spatial intersections with overlap areas, EPSG:5070",
  "main.ak_tract10_3338" = "Alaska census tracts (2010), EPSG:3338",
  "main.ak_block10_group_3338" = "Alaska census block groups (2010), EPSG:3338",
  "main.ak_epa_water_v3_3338" = "Alaska EPA v3 water system boundaries, EPSG:3338",
  "main.ak_tract10_pws_v3_3338" = "Alaska tract × water system spatial intersections, EPSG:3338",
  "main.ak_block10_group_pws_v3_3338" = "Alaska block group × water system spatial intersections, EPSG:3338",
  "main.hi_tract10_6628" = "Hawaii census tracts (2010), EPSG:6628",
  "main.hi_block10_group_6628" = "Hawaii census block groups (2010), EPSG:6628",
  "main.hi_epa_water_v3_6628" = "Hawaii EPA v3 water system boundaries, EPSG:6628",
  "main.hi_tract10_pws_v3_6628" = "Hawaii tract × water system spatial intersections, EPSG:6628",
  "main.hi_block10_group_pws_v3_6628" = "Hawaii block group × water system spatial intersections, EPSG:6628",
  "weighting.us_tract10_pws_v3" = "US-wide tract × water system intersections, all regions combined (no geometry)",
  "weighting.us_bg10_group_pws_v3" = "US-wide block group × water system intersections, all regions combined (no geometry)"
)

raw_counts <- dbGetQuery(
  con,
  "
  SELECT
    schema_name AS table_schema,
    table_name,
    estimated_size AS row_count
  FROM duckdb_tables()
  ORDER BY schema_name, table_name
"
)

overview <- raw_counts |>
  mutate(
    table_key = paste0(table_schema, ".", table_name),
    Description = table_descriptions[table_key],
    Schema = table_schema,
    Table = table_name,
    `Row Count` = format(row_count, big.mark = ",")
  ) |>
  select(Schema, Table, `Row Count`, Description)

reactable(
  overview,
  columns = list(
    Schema = colDef(
      minWidth = 80,
      style = list(fontFamily = "monospace", fontSize = "0.85em")
    ),
    Table = colDef(
      minWidth = 200,
      style = list(fontFamily = "monospace", fontSize = "0.85em")
    ),
    `Row Count` = colDef(minWidth = 90, align = "right"),
    Description = colDef(minWidth = 300)
  ),
  groupBy = "Schema",
  striped = TRUE,
  highlight = TRUE,
  bordered = TRUE,
  pagination = FALSE,
  theme = reactableTheme(
    headerStyle = list(background = "#f1f5f9", fontWeight = "600")
  )
)
```

### Column Detail

```{r}
#| label: 'column-detail'
columns_df <- dbGetQuery(
  con,
  "
  SELECT
    table_schema,
    table_name,
    column_name,
    data_type,
    ordinal_position
  FROM information_schema.columns
  ORDER BY table_schema, table_name, ordinal_position
"
)

col_display <- columns_df |>
  mutate(
    Schema = table_schema,
    Table = table_name,
    Column = column_name,
    Type = data_type
  ) |>
  select(Schema, Table, Column, Type)

reactable(
  col_display,
  columns = list(
    Schema = colDef(
      minWidth = 80,
      style = list(fontFamily = "monospace", fontSize = "0.85em")
    ),
    Table = colDef(
      minWidth = 200,
      style = list(fontFamily = "monospace", fontSize = "0.85em")
    ),
    Column = colDef(
      minWidth = 180,
      style = list(fontFamily = "monospace", fontSize = "0.85em")
    ),
    Type = colDef(
      minWidth = 100,
      style = list(fontFamily = "monospace", fontSize = "0.85em")
    )
  ),
  groupBy = c("Schema", "Table"),
  striped = TRUE,
  highlight = TRUE,
  bordered = TRUE,
  searchable = TRUE,
  pagination = FALSE,
  theme = reactableTheme(
    headerStyle = list(background = "#f1f5f9", fontWeight = "600")
  )
)
```

```{r}
#| label: 'db-disconnect'
#| include: false
dbDisconnect(con)
```
